---
title: "Figure 4"
output: html_document
date: "2024-03-20"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r setup, include=FALSE}
# TCR-seq analysis main folder: /common/bermanblab/data/private_data/POC_ROC/POC_ROC_RNA-Seq/TCR_RNA_seq

# results for TRUST4 are here which were used in the paper: /common/bermanblab/data/private_data/POC_ROC/POC_ROC_RNA-Seq/TCR_RNA_seq/TRUST4

# results for Bowtell data were generated here: /common/bermanblab/data/private_data/POC_ROC/POC_ROC_RNA-Seq/TCR_RNA_seq/TRUST4/bowtell_TCR_seq

# UMAP was generated on the GPU machine:
#How to login to Cerberus:
#ssh kruttika@10.226.39.63
#Password: junkdnalab

# for making the UMAPs, I tried multiple methods to generate it but the final results are here on the GPU machine: /local-data/kruttika/umap_akoya/immune_files/umap_immune.py (parameters were changed here to do scaling with max_value = 10, only immune cell markers were used and then normalized with harmony)


```

# load libraries
```{r message=FALSE}
rm(list = ls(all.names = TRUE))
gc()

library(readxl)
library(tidyverse)
library(ComplexHeatmap)
library(ggplot2)
library(ggpubr)
library(ggrepel)
library(factoextra)

```

# code for how the clonotype counts and RNA-seq data were combined - this chunk of code was not run
```{r message=FALSE, eval=FALSE}

clonotype_count <- read_excel("./Data_for_paper_figures/Clonotype_count_POCROC.xlsx")
colnames(clonotype_count) <- c("RNA_sample", "Clonotype_count", "sample")

clonotype_count$temp <- clonotype_count$RNA_sample

clonotype_count <- separate(clonotype_count, col = temp, into = c("sample", "TCR"), sep = "\\.")
clonotype_count <- separate(clonotype_count, col = sample, into = c("Patient_ID", "Tissue_ID", "site"), sep = "-")
clonotype_count <- separate(clonotype_count, col = Patient_ID, into = c("Patient_ID", "Tissue"), sep = "_")
clonotype_count$Tissue_ID <- ifelse(is.na(clonotype_count$Tissue_ID) == TRUE, clonotype_count$Tissue,
                                    clonotype_count$Tissue_ID)

#TCGA, RtOv, LN
clonotype_count$Tissue_ID <- ifelse(clonotype_count$Tissue_ID == "TCGA", 
                                    clonotype_count$Tissue, clonotype_count$Tissue_ID)
clonotype_count$Tissue_ID <- ifelse(clonotype_count$Tissue_ID == "RtOv", 
                                    clonotype_count$Tissue, clonotype_count$Tissue_ID)
clonotype_count$Tissue_ID <- ifelse(clonotype_count$Tissue_ID == "LN", 
                                    clonotype_count$Tissue, clonotype_count$Tissue_ID)
clonotype_count$Tissue_ID2 <- paste(clonotype_count$Tissue_ID, clonotype_count$site, sep = "-")
clonotype_count$Tissue_ID2 <- ifelse(is.na(clonotype_count$site) == TRUE, clonotype_count$Tissue_ID,
                                     clonotype_count$Tissue_ID2)


clonotype_count <- subset(clonotype_count, select = -c(Tissue, Tissue_ID, site))
colnames(clonotype_count) <- c(colnames(clonotype_count)[-5], "Tissue_ID")

image_results <- read.table("./Data_for_paper_figures/all_image_results_011623.txt",
                            sep = "\t", header = T)

setdiff(clonotype_count$Tissue_ID, image_results$Tissue_ID)

image_results$Tissue_ID[grep("4601", image_results$Tissue_ID)] <- "4601"
image_results <- mutate(image_results, Patient_ID = as.character(Patient_ID))

clonotype_count <- left_join(clonotype_count, image_results) 
sum(is.na(clonotype_count))
clonotype_count <- na.omit(clonotype_count)
# do not have image coordinates for 5105


good.samples <- read.table("./Data_for_paper_figures/good.samples.tsv",
                           sep = "\t", header = T)

# always check: 21020, 22720-2547, 24892, 25236, 26862-4860, 30961, 48986
good.samples$HRD_status[grep("48986", good.samples$patientID)] <- c("HRP", "HRP")
good.samples$HRD_status[grep("2547", good.samples$tissueID)] <- "HRP"

setdiff(good.samples$tissueID, clonotype_count$Tissue_ID)
setdiff(clonotype_count$Tissue_ID, good.samples$tissueID)

clonotype_count$Tissue_ID[grep("3209", clonotype_count$Tissue_ID)] <- "3209"
clonotype_count$Tissue_ID[grep("1624", clonotype_count$Tissue_ID)] <- "1624"

good.samples$tissueID[grep("1671", good.samples$tissueID)] <- c("1671-RtOv", "1671-Omm")
good.samples$tissueID[grep("4416", good.samples$tissueID)] <- c("4416-RtLN", "4416-LtLN")

clonotype_count$Tissue_ID[grep("2604", clonotype_count$Tissue_ID)] <- "2604"

colnames(good.samples) <- c(colnames(good.samples)[1], "Patient_ID", "Tissue_ID",
                            colnames(good.samples)[4:20])
good.samples <- mutate(good.samples, Patient_ID = as.character(Patient_ID)) %>% 
  mutate(., Tissue_ID = as.character(Tissue_ID))

clonotype_count <- left_join(clonotype_count, good.samples, by = c("Patient_ID", "Tissue_ID"))

sum(is.na(clonotype_count))

# saving the RNA-seq data with imaging info, clonotype count and epitope count
setdiff(clonotype_count$Tissue_ID, epitope_count$Tissue_ID)

# 4601 and 2894 not processed for epitope count

epitope_count$Tissue_ID[grep("3209", epitope_count$Tissue_ID)] <- "3209"
clonotype_count <- left_join(clonotype_count, epitope_count, by = c("Patient_ID", "Tissue_ID"))
clonotype_count <- clonotype_count[,-c(4,13:17,23,25:30,33)]
sum(is.na(clonotype_count))

#write.table(clonotype_count, "clonotype_image_epitope_RNA_no5105.txt",
 #           sep = "\t", quote = F, col.names = T, row.names = F)
```

# Fig.4.A; Correlation hovernet and TCR-seq (n = 40) 
```{r message=FALSE}

library(ggplot2)
image_clonotype <- read.table("./Data_for_paper_figures/clonotype_image_epitope_RNA_no5105.txt",
                              sep = "\t", header = T)


image_clonotype <- image_clonotype[-grep("1624",image_clonotype$Tissue_ID),]
image_clonotype <- image_clonotype[-grep("1671-Omm",image_clonotype$Tissue_ID),]
image_clonotype <- image_clonotype[-grep("4691",image_clonotype$Tissue_ID),]
image_clonotype <- image_clonotype[-grep("3322",image_clonotype$Tissue_ID),]
image_clonotype <- image_clonotype[-grep("4513",image_clonotype$Tissue_ID),]
image_clonotype <- image_clonotype[-grep("3360",image_clonotype$Tissue_ID),]
image_clonotype <- image_clonotype[-grep("4612",image_clonotype$Tissue_ID),]
image_clonotype <- image_clonotype[-grep("4448",image_clonotype$Tissue_ID),]
image_clonotype <- image_clonotype[-grep("4704",image_clonotype$Tissue_ID),]

cor.test(image_clonotype$Clonotype_count, image_clonotype$Necrosis) # r=0.22, p-value = 0.18
cor.test(image_clonotype$Clonotype_count, image_clonotype$Inflammatory) # r=0.49, p-value = 0.0014
cor.test(image_clonotype$Clonotype_count, image_clonotype$Connective) # r=0.21, p-value = 0.19
cor.test(image_clonotype$Clonotype_count, image_clonotype$Neoplastic) # r=-0.45, p-value = 0.0036

p1 <- ggplot(image_clonotype, aes(x = Clonotype_count, y = Inflammatory, col = HRD_status)) +
  geom_point(size = 4) +
  theme_bw() +
  theme(text = element_text(size = 25),
        legend.position = "None",
        aspect.ratio = 1) +
  scale_color_manual(values = c("#D53E4F", "#3288BD")) +
  labs(x = "", y = "% of Inflammatory cells",
       col = "HRD status") +
  geom_smooth(method=lm , color="black", fill="grey", se=TRUE)


p2 <- ggplot(image_clonotype, aes(x = Clonotype_count, y = Neoplastic, col = HRD_status)) +
  geom_point(size = 4) +
  theme_bw() +
  theme(text = element_text(size = 25),
        legend.position = "None",
        aspect.ratio = 1) +
  scale_color_manual(values = c("#D53E4F", "#3288BD")) +
  labs(x = "", y = "% of Neoplastic cells",
       col = "HRD status") +
  geom_smooth(method=lm , color="black", fill="grey", se=TRUE)

p3 <- ggplot(image_clonotype, aes(x = Clonotype_count, y = Connective, col = HRD_status)) +
  geom_point(size = 4) +
  theme_bw() +
  theme(text = element_text(size = 25),
        legend.position = "None",
        aspect.ratio = 1) +
  scale_color_manual(values = c("#D53E4F", "#3288BD")) +
  labs(x = "Clonotype count", y = "% of Connective cells",
       col = "HRD status") +
  geom_smooth(method=lm , color="black", fill="grey", se=TRUE)

p4 <- ggplot(image_clonotype, aes(x = Clonotype_count, y = Necrosis, col = HRD_status)) +
  geom_point(size = 4) +
  theme_bw() +
  theme(text = element_text(size = 25),
        legend.position = "None",
        aspect.ratio = 1) +
  scale_color_manual(values = c("#D53E4F", "#3288BD")) +
  labs(x = "Clonotype count", y = "% of Necrotic cells",
       col = "HRD status") +
  geom_smooth(method=lm , color="black", fill="grey", se=TRUE)


#png("./figures/clonotype_inflamm.png", units = "in", height = 5, width = 5, res = 600)
p1
#dev.off()

#png("./figures/clonotype_neoplastic.png", units = "in", height = 5, width = 5, res = 600)
p2
#dev.off()

#png("./figures/clonotype_connective.png", units = "in", height = 5, width = 5, res = 600)
p3
#dev.off()

#png("./figures/clonotype_necrotic.png", units = "in", height = 5, width = 5, res = 600)
p4
#dev.off()
```

# Fig.4.B-left panel; Morrisita horn index, point plot 
```{r message=FALSE}
morisita_idx_pocroc <- read.table("./Data_for_paper_figures/morisita_index_overlap.txt",
                                  sep = "\t", header = T)
morisita_idx_pocroc <- na.omit(morisita_idx_pocroc)

my_comparisons <- list( c("HRD", "HRP"))


morisita_idx_pocroc$log2_idx <- log2(morisita_idx_pocroc$morisita_idx+1)

#png("./figures/TCR-seq_morisita_idx_POCROC.png", units = "in", height = 5, 
    #width = 4, res = 600)
ggplot(morisita_idx_pocroc, aes(x = HRD_status, y = morisita_idx, col = HRD_status)) +
  geom_point(position = position_jitter(width = 0.15)) +
  scale_color_manual(values = c("#AA323F", "#286D97")) +
  theme_bw() +
  theme(legend.position = 'None',
        text = element_text(size = 25)) +
  labs(y = "Morisita-Horn Index") +
  ylim(c(-0.01,0.08))
#dev.off()

##### not run this code below

#p <- ggviolin(morisita_idx_pocroc, x = "HRD_status", y = "morisita_idx", 
 #             fill = "HRD_status", add = "boxplot", width = 1.5,
  #            add.params = list(fill = NA, width = 0.05),
   #           palette = c("#AA323F", "#286D97"),
    #          ylab = "Morisita-Horn index")
# Use only p.format as label. Remove method name.
#png("./figures/TCR-seq_morisita_idx_POCROC.png", units = "in", height = 7, 
 #   width = 6, res = 600)
#p + geom_jitter(shape=16, position=position_jitter(0.02)) +
 # stat_compare_means(comparisons = my_comparisons, 
  #                   method = "t.test",  size = 8) +
  #theme(text = element_text(size = 25),
   #     axis.title.x = element_blank(),
    #    axis.text.x = element_text(angle = 45, hjust = 1),
     #   legend.position = "None",
      #  panel.border = element_rect(colour = "black", fill=NA, size=0.5)) +
  #scale_y_continuous(breaks = seq(0, 0.4, by = 0.04)) +
  #labs(fill = "")
#dev.off()
```

# Fig.4.B-right panel; clonotype count overlap
```{r message=FALSE}


clonotype_overlap_pocroc <- read.table("./Data_for_paper_figures/clonotype_count_overlap.txt",
                                       sep = "\t", header = T)
clonotype_overlap_pocroc <- na.omit(clonotype_overlap_pocroc)
my_comparisons <- list( c("HRD", "HRP"),
                        c("HRD", "HRD-HRP"),
                        c("HRP", "HRD-HRP"))



p2 <- ggviolin(clonotype_overlap_pocroc, x = "group", fill = "group",
               y = "clonotype_count", add = "boxplot", width = 1.15,
               add.params = list(fill = NA, width = 0.05),
               palette = c("#AA323F", "#286D97", "black"),
               ylab = "# of shared clonotypes")
# Use only p.format as label. Remove method name.
#png("./figures/TCR-seq_shared_clonotype_POCROC.png", units = "in", 
 #   height = 8, width = 6, res = 600)
p2 + stat_compare_means(comparisons = my_comparisons, 
                        method = "t.test",  size = 8) +
  theme(text = element_text(size = 25),
        axis.title.x = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "None",
        panel.border = element_rect(colour = "black", fill=NA, size=0.5)) +
  labs(fill = "") +
  scale_y_continuous(breaks = seq(0, 7, by = 1))
#dev.off()
```

# Fig.4.C; immunarch shared clonotype counts HRD and HRP
```{r message=FALSE}
library(immunarch)
library(assertthat)
library(ggplot2)
library(ggseqlogo)

immdata_pocroc_trust4 <- repLoad("./immunarch_input_TRUST4_TRB/correct_format")



imm_ov1 <- repOverlap(immdata_pocroc_trust4$data, .method = "public", .verbose = F)

imm_ov1 <- as.data.frame(imm_ov1)
imm_ov1[lower.tri(imm_ov1)] <- NA

imm_ov1 <- rownames_to_column(imm_ov1, "pair1")

imm_ov1_gather <- gather(imm_ov1, key = "pair2", value = "count", -pair1)
imm_ov1_gather <- na.omit(imm_ov1_gather)

# removing tissues with <= 1 clonotype identified
imm_ov1_gather <- imm_ov1_gather[-grep("1624", imm_ov1_gather$pair1),]
imm_ov1_gather <- imm_ov1_gather[-grep("2112", imm_ov1_gather$pair1),]
imm_ov1_gather <- imm_ov1_gather[-grep("2883", imm_ov1_gather$pair1),]
imm_ov1_gather <- imm_ov1_gather[-grep("2722", imm_ov1_gather$pair1),]


imm_ov1_gather <- separate(imm_ov1_gather, col = pair1, into = c("patient", "random"),
                           sep = "\\.")

imm_ov1_gather <- separate(imm_ov1_gather, col = pair2, into = c("patient2", "random"),
                           sep = "\\.")


pair1 <- substring(imm_ov1_gather$patient, 7)
pair2 <- substring(imm_ov1_gather$patient2, 7)


#write.csv(unique(c(pair1, pair2)), "temp_tissue.csv")

tissueID_HRD <- read.csv("./Data_for_paper_figures/temp_tissue.csv")

imm_ov1_gather$pair1 <- pair1
imm_ov1_gather$pair2 <- pair2


pair1 <- substr(imm_ov1_gather$patient, 1,5)
pair2 <- substr(imm_ov1_gather$patient2, 1,5)

imm_ov1_gather$pair1_ID <- pair1 
imm_ov1_gather$pair2_ID <- pair2 

imm_ov1_gather_HRD <- filter(imm_ov1_gather, pair1 %in% (filter(tissueID_HRD, HRD_status == "HRD"))$tissueID & pair2 %in% (filter(tissueID_HRD, HRD_status == "HRD"))$tissueID)


imm_ov1_gather_HRD$patient_cat <- "Interpatient"
imm_ov1_gather_HRD$patient_cat[which(imm_ov1_gather_HRD$pair1_ID == imm_ov1_gather_HRD$pair2_ID)] <- "Intrapatient"

imm_ov1_gather_HRD <- subset(imm_ov1_gather_HRD, select = c(count,pair1_ID, patient_cat))

colnames(imm_ov1_gather_HRD) <- c("overlap_count", "PatientID", "patient_cat")


imm_ov1_gather_HRP <- filter(imm_ov1_gather, pair1 %in% (filter(tissueID_HRD, HRD_status == "HRP"))$tissueID &
                               pair2 %in% (filter(tissueID_HRD, HRD_status == "HRP"))$tissueID)

imm_ov1_gather_HRP$patient_cat <- "Interpatient"
imm_ov1_gather_HRP$patient_cat[which(imm_ov1_gather_HRP$pair1_ID == imm_ov1_gather_HRP$pair2_ID)] <- "Intrapatient"

imm_ov1_gather_HRP <- subset(imm_ov1_gather_HRP, select = c(count,pair1_ID, patient_cat))

colnames(imm_ov1_gather_HRP) <- c("overlap_count", "PatientID", "patient_cat")

#png("HRD_TCR_inter_intra.png", units = "in", height = 5, 
 #3   width = 5, res = 600)
ggplot(imm_ov1_gather_HRD, aes(x = patient_cat, y = overlap_count, color = patient_cat)) +
  geom_point(size = 2.5,position = position_jitter(width = 0.15)) +
  scale_color_manual(values = c('grey', "#fdae61")) +
  theme_bw() +
  ylim(c(-1.5,6)) + theme(legend.position = "none",
                          text = element_text(size = 25)) +
  labs(y="# shared clonotypes")
#dev.off()

#png("HRP_TCR_inter_intra.png", units = "in", height = 5, 
 #   width = 5, res = 600)
ggplot(imm_ov1_gather_HRP, aes(x = patient_cat, y = overlap_count, color = patient_cat)) +
  geom_point(size = 2.5,position = position_jitter(width = 0.15)) +
  scale_color_manual(values = c('grey', "#fdae61")) +
  theme_bw() +
  ylim(c(-1.5,6)) + theme(legend.position = "none",
                          text = element_text(size = 25)) +
  labs(y="# shared clonotypes")
#dev.off()

t.test(imm_ov1_gather_HRD$overlap_count ~ imm_ov1_gather_HRD$patient_cat)
t.test(imm_ov1_gather_HRP$overlap_count ~ imm_ov1_gather_HRP$patient_cat)
```

# Fig.4.D; survival curve clonotype diversity
```{r message=FALSE}
patient_survival <- read.csv("./Data_for_paper_figures/Patient_survival.csv")
colnames(patient_survival) <- c("patientID", "time_to_death")
good.samples <- read.table("./Data_for_paper_figures/good.samples.tsv",
                           sep = "\t", header = T)

good.samples$HRD_status[grep("48986", good.samples$patientID)] <- c("HRP", "HRP")
good.samples$HRD_status[grep("22720", good.samples$patientID)] <- c("HRP", "HRP")

clonotype_count_pocroc <- read.table("./Data_for_paper_figures/clonotype_count_POCROC.txt",
                                     sep = "\t", header = T)
metadata_pocroc <- read.table("./immunarch_input_TRUST4_TRB/correct_format/metadata.txt", sep = "\t", header = T)

clonotype_count_pocroc <- left_join(clonotype_count_pocroc, metadata_pocroc)
clonotype_count_pocroc <- left_join(clonotype_count_pocroc, good.samples)

div_gini <- read.table("./Data_for_paper_figures/gini_index_POCROC.txt", sep = "\t", header = T)
setdiff(div_gini$Sample, clonotype_count_pocroc$Sample)
colnames(div_gini) <- c("Sample", "Gini_index")

clonotype_count_pocroc <- left_join(clonotype_count_pocroc, div_gini)

clonotype_count_mean <-  clonotype_count_pocroc %>% group_by(patientID) %>% 
  summarise(Mean_count = mean(Gini_index))

clonotype_count_mean$clonotype_high_low <- ifelse(clonotype_count_mean$Mean_count < summary(clonotype_count_mean$Mean_count)[[2]],
                                                  "Low clonotype diversity", "High clonotype diversity")

clonotype_count_mean <- left_join(clonotype_count_mean, patient_survival)
clonotype_count_mean$censored <- 1
clonotype_count_mean$censored[grep("Alive", clonotype_count_mean$time_to_death)] <- 0

clonotype_count_mean$time_to_death[grep("0", clonotype_count_mean$censored)] <- "7300"


clonotype_count_mean[,grep("death", colnames(clonotype_count_mean))] <- lapply(grep("death", colnames(clonotype_count_mean)), 
                                                                               function(x) as.numeric(clonotype_count_mean[[x]]))

clonotype_count_mean$time_to_death <- clonotype_count_mean$time_to_death/365


library(survminer)
library(survival)
surv_object <- Surv(time = clonotype_count_mean$time_to_death,
                    event = clonotype_count_mean$censored)
fit1 <- survfit(surv_object ~ clonotype_high_low, data = clonotype_count_mean)

#png("./figures/TCR-seq_surv_avg_gini_index.png", units = "in", 
 #   height = 8, width = 10, res = 600)
ggsurvplot(fit1, data = clonotype_count_mean, pval = TRUE, risk.table = TRUE,
           palette = c("#5CBD9F", "#5e4fa2"), 
           font.tickslab = c(20),
           legend.labs=c("High clonotype diversity", "Low clonotype diversity"),
           legend.title = "",
           ggtheme = theme_classic2(base_size=20, base_family = "Arial"),
           font.family = "Arial") +
  labs(x = "Time (days)")
#dev.off()

# mean/sum clonotype count with low considered as below first quandrant p value 0.06 for survival
#mean gini index with low considered as below first quandrant p value 0.029 for survival
```

# Fig.4.E- left panel; epitope count POCROC 
```{r message=FALSE}
epitope_count <- read_excel("./Data_for_paper_figures/epitope_count_011623.xlsx")

epitope_count$prim_recur <- "Recurrent"

epitope_count$prim_recur[grep("Primary", epitope_count$Tumor)] <- "Primary"

epitope_count$prim_HRD <- paste(epitope_count$prim_recur, epitope_count$HRD, sep = "-")

my_comparisons <- list( c("Primary-HRD", "Recurrent-HRD"),
                        c("Primary-HRP", "Recurrent-HRP"),
                        c("Primary-HRD", "Primary-HRP"),
                        c("Recurrent-HRD", "Recurrent-HRP"))

epitope_count <- epitope_count[order(epitope_count$prim_HRD),]

p <- ggboxplot(epitope_count, x = "prim_HRD", y = "epitope count",
               fill = "prim_HRD", palette = c("#AA323F",'#286D97', "#E99DA5",'#96C2DD'),
               add = "jitter",
               ylab = "Epitope count")

#png("./figures/epitope_count_prim_HRD.png", units = "in", height = 10, 
 #   width = 12, res = 600)
p + stat_compare_means(comparisons = c(my_comparisons), 
                       method = "t.test",  size = 8) +
  theme(text = element_text(size = 25),
        axis.text.x=element_blank(),
        axis.ticks = element_blank(),
        axis.title.x = element_blank(),
        panel.border = element_rect(colour = "black", fill=NA, size=0.5),
        legend.position = "top") +
  labs(fill = "")
#dev.off()
```

# Fig.4.E- right panel; epitope count PCAWG
```{r message=FALSE}
pcawg_neo <- read.table("./Data_for_paper_figures/TestPCAWG.txt", sep = "\t", header = T)

table(pcawg_neo$HRD)

HRD_colors <- c("#AA323F",'#286D97')

library(ggpubr)
comparison_HRD <- list(c("HRD", "HRP"))

p <- ggboxplot(pcawg_neo, x = "HRD", y = "epitope.count",
               fill = "HRD", palette = HRD_colors,
               add = "jitter",
               ylab = "PCAWG Epitope Count")

#png("./figures/PCAWG_neoantigen_HRD.png", units = "in", height = 6, 
 #   width = 6, res = 600)
p + stat_compare_means(comparisons = comparison_HRD, 
                       method = "wilcox.test",  size = 8, ylab = 2400) +
  theme(text = element_text(size = 15),
        axis.text.x=element_blank(),
        axis.ticks = element_blank(),
        axis.title.x = element_blank(),
        legend.position = "top") +
  labs(fill = "") + ylim(c(200, 2500))
#dev.off()
```

# loading proteomics data
```{r message=FALSE}
source("Read_proteomics_data.R")

```

# Figure 4.F - ERAP1 correlation with epitope count
```{r message=FALSE}
epitope_count <- read_excel("./Data_for_paper_figures/epitope_count_011623.xlsx")

epitope_count$prim_recur <- "Recurrent"

epitope_count$prim_recur[grep("Primary", epitope_count$Tumor)] <- "Primary"

epitope_count$prim_HRD <- paste(epitope_count$prim_recur, epitope_count$HRD, sep = "-")

coldata_prot_tumor <- filter(coldata_prot, re_classify == "Tumor")

setdiff(coldata_prot_tumor$Tissue_ID, epitope_count$Tissue_ID)

coldata_prot_tumor$Tissue_ID[grep("1624", coldata_prot_tumor$Tissue_ID)] <- "1624-D"
coldata_prot_tumor$Tissue_ID[grep("2604", coldata_prot_tumor$Tissue_ID)] <- c("2604-RtOv", "2604-LtOv") 
coldata_prot_tumor$Tissue_ID[grep("4416", coldata_prot_tumor$Tissue_ID)] <- c("4416-RtLN") 
coldata_prot_tumor$Tissue_ID[grep("1671", coldata_prot_tumor$Tissue_ID)] <- c("1671-Omm", "1671-RtOv") 
coldata_prot_tumor$Tissue_ID[grep("2105", coldata_prot_tumor$Tissue_ID)] <- "2105-C"
coldata_prot_tumor$Tissue_ID[grep("4336", coldata_prot_tumor$Tissue_ID)] <- "4336-B"
coldata_prot_tumor$Tissue_ID[grep("4612", coldata_prot_tumor$Tissue_ID)] <- "4612-B"
coldata_prot_tumor$Tissue_ID[grep("4077", coldata_prot_tumor$Tissue_ID)] <- "4077-B"
coldata_prot_tumor$Tissue_ID[grep("3209", coldata_prot_tumor$Tissue_ID)] <- "3209-TCGA-B"
coldata_prot_tumor$Tissue_ID[grep("4860", coldata_prot_tumor$Tissue_ID)] <- "4860-B"
coldata_prot_tumor$Tissue_ID[grep("4315", coldata_prot_tumor$Tissue_ID)] <- "4315-LN"
coldata_prot_tumor$Tissue_ID[grep("2562", coldata_prot_tumor$Tissue_ID)] <- c("2562-RtOv", "2562-RtOv") 

coldata_prot_tumor <- rownames_to_column(coldata_prot_tumor, "sample")
coldata_prot_tumor <- left_join(coldata_prot_tumor, epitope_count)

sum(is.na(coldata_prot_tumor))

coldata_prot_tumor <- na.omit(coldata_prot_tumor) # 4601 epitope count not quantified

prot_data_erap1 <- as.data.frame(t((filter(prot_data, rownames(prot_data) == "ERAP1_HUMAN"))))

prot_data_erap1 <- rownames_to_column(prot_data_erap1, "sample")

coldata_prot_tumor <- left_join(coldata_prot_tumor, prot_data_erap1)


cor.test(coldata_prot_tumor$ERAP1_HUMAN, coldata_prot_tumor$`epitope count`)

coldata_prot_tumor <- coldata_prot_tumor[order(coldata_prot_tumor$prim_HRD),]

#png("./figures/ERAP1_epitope_count.png", units = "in", height = 5, width = 6, res = 600)
ggplot(coldata_prot_tumor, aes(y = ERAP1_HUMAN, x = `epitope count`, col = prim_HRD)) +
  geom_point(size = 4) +
  theme_bw() +
  theme(text = element_text(size = 15),
        aspect.ratio = 1) +
  scale_color_manual(values = c("#AA323F", '#286D97', "#E99DA5",'#96C2DD')) +
  labs(y = "ERAP1 log2(Protein intensity)", x = "Epitope count",
       col = "") +
  ggtitle("Pearson's correlation = -0.3; p-value = 0.025") +
  geom_smooth(method=lm , color="black", fill="grey", se=TRUE)
#dev.off()
```

# Fig.4.G; heatmap of cell phenotyping provided by Akoya 
```{r message=FALSE, eval=FALSE}
# this was provided by Bassem
```

# Fig.4.G - UMAPs
```{r message=FALSE, eval=FALSE}
## jupyter notebook named:UMAP_Figure4.G.ipynb for plotting the UMAPs
```

# Fig.4H - barplots to go below the UMAPs
```{r message=FALSE}

df <- read.table("./Data_for_paper_figures/merged_cell_summary_akoya.txt", sep = "\t", header = F) 
colnames(df) <- c("cell_type", "raw_count", "norm_count", "accession_number")

metadata <- read_csv("./Data_for_paper_figures/Akoya_sample_info_25slides_072023.csv")
metadata <- as.data.frame(metadata)

setdiff(metadata$accession_number, df$accession_number)
setdiff(df$accession_number, metadata$accession_number)

# S13-32894-G1 is labelled as S13-32894_G1 in the merged file
# changing this specific accession number in the metadata file

metadata[grep("S13-32894-G1", metadata$accession_number),1] <- "S13-32894_G1"
setdiff(metadata$accession_number, df$accession_number)
setdiff(df$accession_number, metadata$accession_number)

df <- left_join(df, metadata)
sum(is.na(df))

### converting normalized counts (cell count/ total number of cells) to percentages
df$norm_count <- df$norm_count*100

group4_comparison <- list( c("Primary-HRD", "Primary-HRP"),
                           c("Recurrent-HRD", "Recurrent-HRP"))
group4_comparison_prim_recur <- list( c("Primary-HRD", "Recurrent-HRD"),
                                      c("Primary-HRP", "Recurrent-HRP"))

group4_comparison_germline <- list( c("Primary-BRCAmut", "Primary-BRCAwt"),
                                    c("Recurrent-BRCAmut", "Recurrent-BRCAwt"))

group2_HRD <- list( c("HRD", "HRP"))
group2_prim_recur <- list( c("Primary", "Recurrent"))


group4_color <- c("#AA323F",'#286D97', "#E99DA5",'#96C2DD')
HRD_color <- c("#AA323F",'#286D97')
recur_color <- c("#f97c03",'#5e4fa2')

df_immune <- filter(df, !cell_type %in% c("Endothelial cells", "Fibroblasts", "Necrotic Tumor cells", "other",
                                          "Smooth muscle cells", "Tumor cells", "Epithelial cells"))

sum_celltype <- df_immune %>% group_by(cell_type) %>% summarize(sum_celltype = sum(raw_count))
sum_celltype <- as.data.frame(sum_celltype)

sum_celltype$ratios <- sum_celltype$sum_celltype/sum(sum_celltype$sum_celltype)
sum_celltype$group <- "random"

cell_type_col <-  c("B cells" ="#1f77b4", "Cytotoxic T cells" ="#ff7f0e","Dendritic cells"="#279e68",
                    "Helper T cells"= "#d62728","M1 Macrophages"= "#17becf","M2 Macrophages" ="#8c564b",
                    "Monocytes"= "#e377c2", "NK cells"="#b5bd61", 
                    "T cells double negative"="#aec7e8", "Tregs"="#aa40fc")


#png("immune_cells_total_barplot.png", units = "in", height = 12, width = 4, res = 300)
ggplot(sum_celltype, aes(x = group, y = ratios, fill = cell_type)) + 
  geom_col() +
  scale_fill_manual(values = cell_type_col) +
  theme_bw() +
  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
#dev.off()


df$patientID <- substr(df$POCROC_labels, 1,5)
df_immune <- filter(df, !cell_type %in% c("Endothelial cells", "Fibroblasts", "Necrotic Tumor cells", "other",
                                          "Smooth muscle cells", "Tumor cells", "Epithelial cells"))

sum_patientID <- df %>% group_by(patientID) %>% summarize(sum_patientID = sum(raw_count))
sum_patientID <- as.data.frame(sum_patientID)

sum_patientID$ratios <- sum_patientID$sum_patientID/sum(sum_patientID$sum_patientID)
sum_patientID$group <- "random"


cell_type_col <-  c("18975s"="#018700",
                    "16030"= "#d60000","22720"= "#00acc6",
                    "25258"= "#ff7ed1", "24487"="#97ff00", "18915"="#8c3bff",
                    "27561"="#6b004f", "30961"= "#ffa52f",
                    "34001" = "#573b00", "40729" = "#005659",
                    "41245" = "#0000dd", "41323" = "#00fdcf", "45097" = "#a17569")




#png("immune_cells_patientID_total_barplot.png", units = "in", height = 12, width = 4, res = 300)
ggplot(sum_patientID, aes(x = group, y = ratios, fill = patientID)) + 
  geom_col() +
  scale_fill_manual(values = cell_type_col) +
  theme_bw() +
  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
#dev.off()


```

